Comparing ChatGPT-4 to Physicians for Assigning Causes of Death from Verbal Autopsy: A Case Study of 6939 Deaths in Sierra Leone from 2019-2022

### 2024-02-07 manuscript/manuscript.tex
Add methods figure

This study uses 6939 physician agreed records of 11,920 records collected in 2019 to 2022 from the Healthy Sierra Leone (HEAL-SL) study as described in Section \ref{methods-data}. Section \ref{methods-models} describes the four models, ChatGPT-3.5, ChatGPT-4, InterVA-5, and InSilicoVA, for COD assignment, and their inputs and outputs, while section \ref{methods-eval} details the performance evaluation of the four models relative to physicians overall, by ages, and by COD, using population and individual level metrics.

@@

The 6939 physician agreed records were relatively evenly distributed (approximately 40-57\% male and female) for each age group in terms of sex (Figure \ref{fig-data-agesex}) with the most records in the 1-5 year (n=1633, 24\%) age range, and the least records in the 10-14 year (n=135, 2\%) and 7-27 day (n=82, 1\%) age ranges (Figure \ref{fig-data-agerange}). Other 5-year age ranges within 0 days to 69 years were relatively evenly distributed with approximately 4-6\% of the 6939 records each. The 3826 adult records had most records in the 65-69 year (n=575, 15\%), and the least in the 10-14 year (n=135, 4\%) age ranges, while other age ranges were relatively evenly distributed at 7-9\% of all records (Figure \ref{fig-data-agerange-adult}). For adult CODs, Malaria (n=799, 21\%) had the highest number of adult records, and cancers (n=49, 1\%), diabetes mellitus (n=27, \textless 1\%), and suicide (n=3, \textless 1\%) had the lowest number adult of adult records, while other CODs had between 2-11\% of all adult records (Figure \ref{fig-data-cod-adult}). The 2636 child records had most records in the 1-5 year (n=1633, 62\%) age range while all other age ranges were relatively evenly distributed at 12-14\% of all records (Figure \ref{fig-data-agerange-child}). For child CODs, Malaria (n=1382, 52\%) and other infections (n=667, 25\%) had the highest number of child records, while nutritional deficiencies (n=11, \textless 1\%) and congenital anomalies (n=1, \textless 1\%) had the lowest number of records with all other CODs between 3-7\% of all child records. The 477 neonatal records only had two age ranges with most records in the 0-6 day (n=395, 83\%) age range and the rest in the 7-27 day (n=82, 17\%) age range (Figure \ref{fig-data-agerange-neo}). For neonatal CODs, stillbirth (n=172, 36\%) had the highest number of neonatal records, and ill-defined (n=23, 5\%), other (n=5, 1\%), and congenital anomalies (n=2, \textless 1\%) had very low number of records with the rest of the CODs between 15-22\% of all neonatal records (Figure \ref{fig-data-cod-neo}). See Appendices \ref{app-data} to \ref{app-data-neo} for a list of all figures presenting distribution of the 6939 physician agreed records.

>>>>

This study uses 6939 physician agreed records of 11,920 records collected in 2019 to 2022 from the Healthy Sierra Leone (HEAL-SL) study as described in Section \ref{methods-data}. Section \ref{methods-models} describes the four models, ChatGPT-3.5, ChatGPT-4, InterVA-5, and InSilicoVA, for COD assignment, and their inputs and outputs, while section \ref{methods-eval} details the performance evaluation of the four models relative to physicians overall, by ages, and by COD, using population and individual level metrics. Figure \ref{fig-methods} details the methods used in this study.

\begin{figure}[h!]
\centering
\includegraphics[width=0.7\textwidth]{fig-methods.pdf}
\caption{Case study methods. Verbal autopsy records in Sierra Leone between 2019-2022 from the HEAL-SL study were filtered for physician agreed records, and pre-processed for COD assignment by ChatGPT-3.5, ChatGPT-4, InterVA-5 and InSilicoVA models. The output models were evaluated with CSMF Accuracy and PCCC overall, by age group, and by five-year age ranges for each age group.}\label{fig-methods}
\end{figure}

@@

The 6939 physician agreed records were relatively evenly distributed (approximately 40-57\% male and female) for the adult, child, and neonatal age groups in terms of sex (Figure \ref{fig-data-agesex}) with the most records in the 1-5 year (n=1633, 24\%) age range, and the least records in the 10-14 year (n=135, 2\%) and 7-27 day (n=82, 1\%) age ranges (Figure \ref{fig-data-agerange}). The other 15 five-year age ranges within 0 days to 69 years were relatively evenly distributed with approximately 4-6\% of the 6939 records each. The 3826 adult records had most records in the 65-69 year (n=575, 15\%), and the least in the 10-14 year (n=135, 4\%) age ranges, while the other 10 age ranges were relatively evenly distributed at 7-9\% of all records (Figure \ref{fig-data-agerange-adult}). For adult CODs, Malaria (n=799, 21\%) had the highest number of adult records, and cancers (n=49, 1\%), diabetes mellitus (n=27, \textless 1\%), and suicide (n=3, \textless 1\%) had the lowest number adult of adult records, while the other 15 CODs had between 2-11\% of all adult records (Figure \ref{fig-data-cod-adult}). The 2636 child records had most records in the 1-5 year (n=1633, 62\%) age range while the three other age ranges were relatively evenly distributed at 12-14\% of all records (Figure \ref{fig-data-agerange-child}). For child CODs, Malaria (n=1382, 52\%) and other infections (n=667, 25\%) had the highest number of child records, while nutritional deficiencies (n=11, \textless 1\%) and congenital anomalies (n=1, \textless 1\%) had the lowest number of records with the five other CODs between 3-7\% of all child records. The 477 neonatal records had two age ranges with most records in the 0-6 day (n=395, 83\%) age range and the rest in the 7-27 day (n=82, 17\%) age range (Figure \ref{fig-data-agerange-neo}). For neonatal CODs, stillbirth (n=172, 36\%) had the highest number of neonatal records, and ill-defined (n=23, 5\%), other (n=5, 1\%), and congenital anomalies (n=2, \textless 1\%) had the least number of records with the rest of the other three CODs between 15-22\% of all neonatal records (Figure \ref{fig-data-cod-neo}). See Appendices \ref{app-data} to \ref{app-data-neo} for a list of figures \ref{fig-data-agesex} to \ref{fig-data-cod-neo} showing distributions of the 6939 physician agreed records.
